---
title: "Data Visualization and Maps II"
subtitle: "HES 505 Fall 2023: Session 26"
author: "Carolyn Koehn"
execute: 
  eval: true
format: 
  revealjs:
    theme: mytheme.scss
    slide-number: true
    show-slide-number: print
    self-contained: true  
---

```{r}
#| include: false
library(tidycensus)
library(sf)
library(tidyverse)
library(ggmap)
library(maps)
library(units)
library(cartogram)
library(patchwork)
library(tmap)
library(viridis)
library(terra)
censkey <- Sys.getenv("CENSUS_API_KEY")
```

# Objectives {background="#9F281A"}

By the end of today you should be able to:

* Generate complicated plot layouts without additional pre-processing

* Construct a map using `ggplot2` and `tmap`

* Combine vector and raster data in the same map


# Building Choropleth Maps {background="#0033A0"}

## `tidycensus` package

https://walker-data.com/tidycensus/articles/basic-usage.html

## Using `ggplot2`

```{r}
#| echo: true
cty.info <- get_acs(geography = "county", 
                      variables = c(pop="B01003_001", 
                                    medincome = "B19013_001"),
                      survey="acs5",
                      state = c("WA", "OR", "ID", "MT", "WY"),
                      geometry = TRUE, key = censkey, progress_bar=FALSE) %>% 
  select(., -moe) %>% 
  pivot_wider(
    names_from = "variable",
    values_from = "estimate"
  )

p <- ggplot(data=cty.info) +
  geom_sf(mapping=aes(fill=medincome))
```

## Static Maps with `ggplot2`

```{r}
p
```

## Changing aesthetics

```{r}
#| echo: true
p <- ggplot(data=cty.info) +
  geom_sf(mapping=aes(fill=pop), color="white") +
  scale_fill_viridis()

```

## Changing aesthetics

```{r}
p
```

## Adding layers

```{r}
#| echo: true
st <- tigris::states(progress_bar=FALSE) %>% 
  filter(., STUSPS %in% c("WA", "OR", "ID", "MT", "WY"))

p <- ggplot(data=cty.info) +
  geom_sf(mapping=aes(fill=pop), color="white") +
  geom_sf(data=st, fill=NA, color="red") +
  scale_fill_viridis()
```

## Adding layers

```{r}
p
```

## Using `tmap`

```{r}
#| echo: true
pt <- tm_shape(cty.info) + 
  tm_polygons(col = "pop",
              border.col = "white") + 
  tm_legend(outside = TRUE)
```

## Using `tmap`

```{r}
pt
```

## Changing aesthetics

```{r}
#| echo: true
pt <- tm_shape(cty.info) + 
  tm_polygons(col = "pop", n=10,palette=viridis(10),
              border.col = "white") + 
  tm_legend(outside = TRUE)
```

## Changing aesthetics

```{r}
pt
```

## Adding layers

```{r}
#| echo: true

pt <- tm_shape(cty.info) + 
  tm_polygons(col = "pop", n=10,palette=viridis(10),
              border.col = "white") + 
  tm_shape(st) +
  tm_borders("red") +
  tm_legend(outside = TRUE)
```

## Adding layers
```{r}
pt

```

## Themes

```{r}
#| echo: true

p
```

## Themes

```{r}
#| echo: true

p + 
  theme_void()
```

## Themes

```{r}
#| echo: true

p + 
  theme_void() +
  theme(panel.background = element_rect(fill="white", color="black"))
```

## Labels

```{r}
#| echo: true

p +
  labs(title = "County Populations")
```

## Legend customization

```{r}
#| echo: true

ggplot(data=cty.info) +
  geom_sf(mapping=aes(fill=pop), color="white") +
  geom_sf(data=st, fill=NA, color="red") +
  scale_fill_viridis(name = "Population\n(millions)",
                     breaks = seq(from=0, to=3e6, by=1e6),
                     labels = seq(from=0, to=3e6, by=1e6)/1e6)
```

## Legend customization

```{r}
#| echo: true

ggplot(data=cty.info) +
  geom_sf(mapping=aes(fill=pop), color="white") +
  geom_sf(data=st, fill=NA, color="red") +
  scale_fill_viridis(name = "Population\n(millions)",
                     limits = c(0, 3e6),
                     breaks = seq(from=0, to=3e6, by=1e6),
                     labels = seq(from=0, to=3e6, by=1e6)/1e6)
```

# Rasters in `ggplot2`

## Convert raster to tidy format

To visualize rasters in `ggplot2`, they need to be converted to a dataframe.

```{r}
#| echo: true
#| eval: false

fire.haz <- rast("/opt/data/data/assignment01/wildfire_hazard_agg.tif")

fire.haz_df <- as.data.frame(fire.haz)

str(fire.haz_df)
```
```{r}
#| include: false

fire.haz <- rast("C:/Users/carolynkoehn/Documents/HES505_Fall_2024/data/2023/assignment01/wildfire_hazard_agg.tif")

fire.haz_df <- as.data.frame(fire.haz, xy=TRUE)

str(fire.haz_df)
```

## Use `geom_raster`

```{r}
#| echo: true
#| warning: false

ggplot(data = fire.haz_df, aes(x=x, y=y, fill = WHP_ID)) +
  geom_raster()
```

## Coordinate system

```{r}
#| echo: true
#| warning: false

ggplot(data = fire.haz_df, aes(x=x, y=y, fill = WHP_ID)) +
  geom_raster() +
  coord_sf(default_crs = crs(fire.haz))
```

## Layering rasters and vectors

```{r}
#| echo: true
#| warning: true

states <- tigris::states(progress_bar=FALSE) %>% 
  filter(., STUSPS %in% c("WA", "OR", "ID")) %>%
  st_transform(crs = st_crs(fire.haz))

p2 <- ggplot(data = fire.haz_df, aes(x=x, y=y, fill = WHP_ID)) +
  geom_raster() +
  geom_sf(data = states, fill=NA, color="red", inherit.aes = FALSE) +
  coord_sf(default_crs = crs(fire.haz))

p3 <- ggplot() +
  geom_raster(data = fire.haz_df, aes(x=x, y=y, fill = WHP_ID)) +
  geom_sf(data = states, fill=NA, color="red") +
  coord_sf(default_crs = crs(fire.haz))
```

## Layering rasters and vectors

```{r}
p2
```

# Complicated layouts with `patchwork`

## Subplots

```{r}
#| echo: true

p2+p3
```

## Combine legends

```{r}
#| echo: true

p2+p3 + plot_layout(guides="collect")
```

## Change layout

```{r}
#| echo: true

p+p3 + plot_layout(nrow=2)
```

## Map insets

```{r}
#| echo: true

conus <- tigris::states() %>%
  filter(!(DIVISION == 0 | STUSPS %in% c("HI", "AK"))) %>%
  st_transform(st_crs(fire.haz))

inset <- ggplot(data = conus) +
  geom_sf(fill="white") +
  geom_sf(data = filter(conus, STUSPS %in% c("ID", "OR", "WA")),
          fill = "gray70", color="red") +
  theme_void() +
  theme(panel.background = element_rect(fill="white", color="black")) +
  coord_sf()
```

## Map insets

```{r}
#| echo: true

layout <- c(
  patchwork::area(t = 1, l = 1, b = 5, r = 4),
  patchwork::area(t = 1, l = 4, b = 2, r = 5),
  patchwork::area(t = 3, l = 4, b = 5, r = 5)
)

p2 + inset + guide_area() + plot_layout(design=layout, guides='collect')
```

