---
title: "Data Visualization and Maps I"
date: "2024-11-11"
---

```{r slides-videos, echo=FALSE, include=FALSE}
source(here::here("R", "slide-things.R"))
```

We’ve spent the last few weeks learning about operations to compile geographic information into databases for visualization and analysis. Because analysis requires you to know something about your data and because visualization is a great way to explore your data (especially when there’s a lot of it), we’ll turn to that next. For the next few weeks, we’ll be looking at different ways to visualize spatial data and the associated approaches in `R`. Note that this could be an entire course by itself, but hopefully you’ll get enough to get started making publication quality maps by the time we’re done

## Resources

- <i class="fas fa-external-link-square-alt"></i> The [Introduction](https://f0nzie.github.io/dataviz-wilke-2020/introduction.html) and [Visualizing Geospatial Data](https://f0nzie.github.io/dataviz-wilke-2020/geospatial-data.html) chapters [Principles of Figure Design](https://f0nzie.github.io/dataviz-wilke-2020/proportional-ink.html) section in [@wilke2019fundamentals] provide a useful set of general introductions to data visualization principles and practice that is “platform agnostic” (though much of Wilke’s work is done in `R`).

- <i class="fas fa-external-link-square-alt"></i> The [Look at Data](https://f0nzie.github.io/dataviz-wilke-2020/introduction.html) and [Draw Maps](https://f0nzie.github.io/dataviz-wilke-2020/geospatial-data.html) chapters in [@healy2018data] revisits many of the same ideas, but focuses specifically on `R` and `ggplot2`.

- <i class="fas fa-external-link-square-alt"></i> This post on [making maps people want to look at](https://www.esri.com/news/arcuser/0112/make-maps-people-want-to-look-at.html) from ESRI is a nice, concise depiction of some core principles for planning a cartographic project.

- <i class="fas fa-external-link-square-alt"></i> [Making maps with `R`](https://geocompr.robinlovelace.net/adv-map.html) by [@lovelace_geocomputation_2019] introduces the `tmap` package for making nice maps with relatively minimal syntax.


## Objectives

By the end of today you should be able to:

- Describe some basic principles of data visualization

- Extend principles of data visualization to the development of maps

- Distinguish between several common types of spatial data visualization



```{r show-slide-tabs, echo=FALSE, results="asis"}
slide_buttons("/slides/24-slides")
```

<div>

```{=html}
<iframe class="slide-deck" src="/slides/24-slides.html"></iframe>
```

</div>


