---
title: "Data Visualization and Maps II"
date: "2024-11-18"
---

```{r slides-videos, echo=FALSE, include=FALSE}
source(here::here("R", "slide-things.R"))
```

`R` provides a number of different packages for generating plots of your data, but `ggplot2` is probably the most common owing to its ability to use consistent syntax to produce a variety of different graphics. In addition to plots of data and model objects, `ggplot2` can also be used with `sf` objects and some raster datasets to generate publication quality maps. We’ll also take a little more time to understand some of the options for building static maps in `R` and look at a few packages that can help you build publication-quality maps without having to move into a new software.

## Resources

- {{< fa book >}} The [Data Visualization: A Practical Introduction](https://socviz.co/) by @healy2018data provides a lot of examples of Tufte-style graphics built with ggplot2.

- {{< fa arrow-up-right-from-square >}} [Graphic design with ggplot2](https://rstudio-conf-2022.github.io/ggplot2-graphic-design/) is an entire course devoted to making beautiful visualizations with `ggplot2`. If nothing else, check out some of the examples!

- {{< fa arrow-up-right-from-square >}} [The `patchwork` package website](https://patchwork.data-imaginist.com/) provides a lot of examples of building complicated layouts with `ggplot2` objects with intuitive syntax.

- {{< fa book >}} The [Drawing Beautiful Maps Programatically with `R`, `sf`, and `ggplot2`](https://r-spatial.org/r/2018/10/25/ggplot2-sf.html) chapter by Mel Moreno and Mathieu Basille provides a nice series of blog posts designed to help you build maps with `sf` objects.

- {{< fa arrow-up-right-from-square >}} [Creating beautiful demographic maps in `R` with the `tidycensus` and `tmap` packages](http://zevross.com/blog/2018/10/02/creating-beautiful-demographic-maps-in-r-with-the-tidycensus-and-tmap-packages/) by Zev Ross illustrates the simplicity of mapping with `tmap`.

- {{< fa arrow-up-right-from-square >}} [Displaying time series, spatial and space-time data with `R`](https://oscarperpinan.github.io/bookvis/) by Oscar Perpiñán Lamigueiro has a bunch of interesting code for producing maps contained in the book by the same name.

## Objectives

By the end of today you should be able to:

- Understand the relationship between the Grammar of Graphics and ggplot syntax

- Describe the various options for customizing ggplots and their syntactic conventions

- Generate complicated plot layouts without additional pre-processing

- Construct a map using ggplot2 and tmap

- Combine vector and raster data in the same map

<!--

```{r show-slide-tabs, echo=FALSE, results="asis"}
slide_buttons("/slides/26-slides")
```

<div>

```{=html}
<iframe class="slide-deck" src="/slides/26-slides.html"></iframe>
```

</div>
-->
